---
title: "Data Science"
output: html_document
---

##Filtering Rows and Columns

In SQL you can filter columns by using SELECT and rows by using WHERE.Let's look at an example using the Lahman baseball database. First we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA'and yearID=1927"
sqldf(query)

```

###More Examples of SQL


If one wants to find the specific codes of words for Batting they would use 'colnames(Batting)'. In these examples, SQL is being used. First one needs to load these packages:

```{r message=FALSE, warning=FALSE}

library(Lahman)

library(sqldf)

```

Find all instances where the Yankees have hit fourty or more homeruns.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and HR>=40"


sqldf(query)

```


Find all instances where a Yankee had more than fourty homeruns but less than sixty strikeouts.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"


sqldf(query)

```

Find all instances of Phillies hitting more than thirty homeruns in the 1970s.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"


sqldf(query)

```


##Order By

With SQL, we can also order the rows in the output by using ORDER BY. For example, suppose you wanted to find every instance of a player hitting more than fifty homeruns. Suppose you want to have the players with the highest homeruns at the top, you can do this the following way:

```{r message=FALSE,warning=FALSE}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50 ORDER BY HR DESC"
sqldf(query)

```

###More Examples of Order By

Find all instances of a player striking out less than ten times. Make sure each player has atleast 400 at-bats(AB). Order them by the least strike outs to the most.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,SO,AB FROM Batting WHERE SO<10 and AB>=400 ORDER BY SO"
sqldf(query)

```

##Aggregation

Aggregation includes summing up (sum), taking the maximum (max) or minimum (min), or averaging(avg) certain categories by grouping them. One can do this by using GROUP BY, then choosing what they would like to do with the catergory or categories. If one was to also use ORDER BY, this function would be listed afterwards in the R Code in the quotations. WHERE is used when filtering categories before you group, however HAVING is used after you group.

###Aggregation Examples

Find the career homerun totals for Babe Ruth.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,sum(HR) FROM Batting WHERE playerID='ruthba01' GROUP BY playerID"
sqldf(query)

```

Find the career homerun totals for all players, but list only those with a total of 600 or more. Order by having greatest total at the top.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,sum(HR) FROM Batting GROUP BY playerID HAVING sum(HR)>=600 ORDER BY sum(HR) DESC"
sqldf(query)

```

What players have the highest average per season homerun totals? List only averages greater than thirty.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,avg(HR) FROM Batting GROUP BY playerID HAVING avg(HR)>30 ORDER BY avg(HR) DESC"
sqldf(query)

```

##Joins

###Examples of Joins

Get the first and last names, along with teamID, yearID, HR, and have the player be Babe Ruth

```{r message=FALSE,warning=FALSE}

query<-"SELECT nameFirst,nameLast,teamID,yearID,HR FROM Batting INNER JOIN Master ON Batting.playerID=Master.playerID WHERE Batting.playerID='ruthba01'"
sqldf(query)

```

Find all instances of players hitting more than fifty homeruns. Include first and last names, the team, the year, and the homeruns.

```{r message=FALSE,warning=FALSE}

query<-"SELECT nameFirst,nameLast,teamID,yearID,HR FROM Batting INNER JOIN Master ON Batting.playerID=Master.playerID WHERE HR>50 ORDER BY HR DESC"
sqldf(query)

```

Babe Ruth, with playerID, team name, yearID, and HR.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,name,Batting.yearID,Batting.HR FROM Batting INNER JOIN Teams ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID WHERE playerID='ruthba01'"
sqldf(query)

```








