---
title: "SQL"
output: html_document
---

##Filtering Rows and Columns:

In SQL you can filter columns by using SELECT and rows by using WHERE.Let's look at an example using the Lahman baseball database. First we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the homerun totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA'and yearID=1927"
sqldf(query)

```

##Baseball Examples:

###SQL:

If one wants to find the specific codes of words for Batting they would use 'colnames(Batting)'. In these examples, SQL is being used. First one needs to load these packages:

```{r message=FALSE, warning=FALSE}

library(Lahman)

library(sqldf)

```

Find all instances where the Yankees have hit fourty or more homeruns.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='NYA' and HR>=40"


sqldf(query)

```


Find all instances where a Yankee had more than fourty homeruns but less than sixty strikeouts.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR,SO FROM Batting WHERE HR>40 and SO<60"


sqldf(query)

```

Find all instances of Phillies hitting more than thirty homeruns in the 1970s.

```{r message=FALSE,warning=FALSE}

query<-"SELECT playerID,teamID,yearID,HR FROM Batting WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"


sqldf(query)

```


###ORDER BY:

With SQL, we can also order the rows in the output. For example, suppose you wanted to find every instance of a player hitting more than fifty homeruns. Suppose you want to have the players with the highest homeruns at the top, you can do this the following way:

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50 ORDER BY HR DESC"
sqldf(query)

```








